\Huge ACM and Climate Change

### 2018-12-12 main.tex
Polish

ACM’s vibrant conference culture has a dark side: it contributes
significantly to climate change.

@@

There is a broad scientific consensus that the earth is warming due to human
emissions of carbon dioxide and other greenhouse gases (Figure
\ref{fig:warming}). This is not a problem we can leave to future
generations: we are up against it now. Indeed, the UN’s Intergovernmental
Panel on Climate Change (IPCC) recently reported that a 40\% decrease in
emissions by 2030 is needed to avoid irreversible damage
(Figure~\ref{fig:IPCC-cover}). Achieving this will require urgent and sustained
commitment at all levels of society, including non-governmental communities
such as universities, companies, and scientific societies like ACM.

Why ACM? Because one of ACM’s main activities is organizing international
conferences, and these produce significant emissions of greenhouse gases,
particularly from air travel. For example, a single round-trip flight from
Philadelphia to Paris emits the equivalent of about 1.7 tons of carbon
dioxide (\COtwoE) per passenger~\cite{carboncalculator}.\iflater\bcp{Trim
  citation?}\fi{} Compare this against the fact that the average US resident
emits about 16.5 tons per year, while the average European emits about 7
tons~\cite{emissions}.  Moreover, these emissions have no near-term
technological fix because jet fuel is difficult to replace with renewable
energy sources~\cite{elec-air}. So reducing the \COtwoE{} footprint of
conferences is going to require a reduction in the footprint of the
attendees’ air travel.

We are the members of an {\em ad hoc} Committee on Climate Change within
SIGPLAN (ACM’s Special Interest Group on Programming Languages) that was
formed in 2016 to consider possible strategies for mitigating the emissions
of our conferences \cite{ClimateCommitteReport}. Among the ones we find most
attractive are: making it easier to use carbon offsets to mitigate aviation
emissions for conferences, choosing conference locations with an eye to
minimizing emissions, co-locating or merging conferences, virtualizing
program committees, and supporting remote participation.

Buying carbon offsets to reduce the impact of air travel is an ``easy win''
\cite{CarbonOFfsetReport}.  The idea of carbon offsets is to pay somebody to plant trees, capture
methane (from landfills or cattle herds, for example), buy cookstoves to
replace inefficient open-fire cooking in poor communities, or carry out
other actions that remove greenhouse gases from the air or prevent their
emission. At the moment, the average air travel to an international
conference can be offset by paying around \$30 \cite{CarbonOFfsetReport}.
ACM could streamline 
the purchase of such offsets and make offsetting the default by folding them
into its conference registration fees.

Unfortunately, carbon offsetting is only a short-term mitigation, since the
potential supply of offsets is far less than the potential demand~\cite{SEI-Report}. We
must also begin significantly reducing emissions, which means reducing the
distance traveled by the sum of all conference participants.

One approach to doing so is to choose conference locations that are close to
the center of mass of their potential participants.  For example, SIGGRAPH
(the Special Interest Group on Graphics) locates in Los Angeles every other
year---an easy call as this is near an international airport and a major hub
of the worldwide entertainment industry. For other research areas, the
effects of location choice are less obvious. For example,
Figure~\ref{fig:confs} plots the 
locations of the four major SIGPLAN conferences over the past five years,
colored according to the estimated average carbon footprint per
participant. The figure shows that there were significant differences across
conferences and across locations---indeed, it is a factor of two between the
lowest and highest. Why is this? One reason might be that the locus of a
conference’s population is closer to one location than another, and changing
locations reduces the total cost. Another might be the presence of a local
population that will only attend if the conference is very close to
them. Our preliminary analysis suggests it may be both. The larger question
is how to use this kind of information to choose the best locations over
time, for a diffuse population.

Another natural idea is to co-locate or consolidate conferences.  This is
already standard practice in some communities---e.g., SIGGRAPH, SC, and CVPR
are all ``the'' conference for their areas.  More communities---for example,
the programming languages and security communities, each of which holds four
main conferences every year---could consider following suit.  A different
twist on this idea would be to hold a single conference in two locations
simultaneously (e.g., Boston and Paris), with a video link between the
sites. Switching to virtual program committee meetings for conferences is
another obvious step. \mwh{data about surprising connection to conference's
  overall footprint?}\bcp{Or leave it for next round of revisions when we
  may have done more/better analysis?}  Indeed, this switch is already
happening: we found in a recent poll of ACM SIG chairs that, across ACM, a
majority of the SIGs’ main conferences now use online PC meetings rather
than in-person ones.

We can attend some conferences online, saving both emissions and travel
time.  Indeed, SIGPLAN already livestreams all its four main conferences,
while SIGACCESS and SIGCHI have experimented with telepresence robots at
their main conferences~\cite{CHI-remote,Neustaedter}. Journal-first
initiatives, which make conference attendance optional, could also help.
Looking further ahead, some communities are already experimenting with fully
virtual 
% \mwh{reality?}
% BCP: Seems redundant with "fully"
conferences~\cite{OS}.  ACM members can contribute to the
development of technologies for connecting people effectively at a distance.

@@

The problem of climate change is too large and too urgent to leave to world
leaders: organizations at every level must play their part in dealing with
their own contributions, raising awareness, and establishing new ways of
doing business in the lower-carbon future that is coming. For ACM, this
especially means developing strategies for reducing the climate impact of
conference-related travel. You can help in a number of ways: 
\begin{itemize}
\item {\bf Get involved} in raising awareness about this issue in the
organizations you are a part of.  For example, ask the organizers of the ACM
conferences you attend what they are doing about reducing or offsetting
emissions from participants’ travel.
\item {\bf Analyze} your organization’s carbon-emitting activities to inform a
discussion of how best to reduce its footprint.
\item {\bf Connect} with others that are engaging with these issues---in
your workplace, in your other communities, and within ACM (e.g., on the
acm-climate mailing list~\cite{acm-climate}).
\end{itemize}

@@

About Crista’s figure:

@@

Also about Crista’s figure:

@@

Also, what about research on telepresence and virtual meetings? There’s obviously a lot of work in this area, but we may need more in order to (a) let us effectively trade off options, e.g. virtual PC meeting vs in-person; (b) improve the state of the art to make remote meetings more effective.
Consider also discussing the diversity implications on travel. Reducing travel expectations might make the field more attractive to women, who may face higher social expectations to stay home (and are more likely to be single parents). Surely there is a good article to cite here. I found https://journals.sagepub.com/doi/abs/10.1177/0950017006066999 but maybe there is something better. I’m looking…

@@

Check where we’ve used ``carbon'' (or CO2) and where ``greenhouse gases''.  Gloss CO2 or CO2e as appropriate.

>>>>

ACM's vibrant conference culture has a dark side: it contributes
significantly to climate change.

@@

A broad scientific consensus warns that the earth is warming due to human
emissions of carbon dioxide and other greenhouse gases (Figure
\ref{fig:warming}). This is not a problem we can leave to future
generations: we are up against it now. Indeed, the UN's Intergovernmental
Panel on Climate Change (IPCC) recently recommended a 40\% decrease in
emissions by 2030 to avoid irreversible damage
(Figure~\ref{fig:IPCC-cover}). Achieving reductions on this scale will
require urgent and sustained commitment at all levels of society, including
non-governmental communities such as universities, companies, and scientific
societies like ACM.

Why ACM? Because one of ACM's main activities is organizing international
conferences, which produce significant emissions of greenhouse gases,
particularly from air travel. For example, a single round-trip flight from
Philadelphia to Paris emits the equivalent of about 1.7 tons of carbon
dioxide (\COtwoE) per passenger~\cite{carboncalculator}\iflater\bcp{Trim
  citation?}\fi---a substantial fraction of the {\em total} yearly emissions
for an average resident of the US (16.5 tons) or Europe (7
tons)~\cite{emissions}.  Moreover, these emissions have no near-term
technological fix because jet fuel is difficult to replace with
renewable energy sources~\cite{elec-air}. 
\iflater\bcp{Trimmed this sentence: Reducing the \COtwoE{} footprint
of conferences will require a reduction in the footprint of the attendees'
air travel.}\fi

We are the members of an {\em ad hoc} Committee on Climate Change within
ACM's Special Interest Group on Programming Languages (SIGPLAN) that was
formed in 2016 to consider possible strategies for mitigating the emissions
of our conferences \cite{ClimateCommitteReport}. Among the ones we find 
attractive are making it easier to use carbon offsets to mitigate aviation
emissions for conferences, choosing conference locations with an eye to
minimizing emissions, co-locating or merging conferences, virtualizing
program committees, and supporting remote participation.

Buying carbon offsets to reduce the impact of air travel
\cite{CarbonOFfsetReport} is an ``easy win'' that conferences should
implement now.  The idea of carbon offsets is to pay somebody to plant
trees, capture methane (from landfills or cattle herds, for example), buy
cookstoves to replace inefficient open-fire cooking in poor communities, or
carry out other actions that remove greenhouse gases from the air or prevent
their emission. At the moment, the average air travel to an international
conference can be offset by paying around \$30 \cite{CarbonOFfsetReport}.
ACM could streamline the purchase of such offsets and make offsetting the
default by folding them into its conference registration fees.

Sadly, carbon offsetting is only a short-term mitigation, since the
potential supply of offsets is far less than the potential
demand~\cite{SEI-Report}. We must also begin significantly reducing
emissions, which means reducing the distance traveled by the sum of all
conference participants.

One way to do this is to choose conference locations that are close to
the center of mass of their likely participants.  For example, the
SIGGRAPH conference on computer graphics
locates in Los Angeles every other
year---an easy call, as this is near an international airport and a major hub
of the worldwide entertainment industry. For other research areas, the
effects of location choice are less obvious. For example,
Figure~\ref{fig:confs} plots the 
locations of the four major SIGPLAN conferences over the past five years,
colored according to the estimated average carbon footprint per
participant. The figure shows significant differences across
conferences and across locations---indeed, a factor of two difference
between the 
lowest and highest. Why is this? One reason might be that the locus of a
conference's population is closer to one location than another, and changing
locations reduces the total cost. Another might be the presence of a local
population that will only attend if the conference is very close to
them. Our preliminary analysis suggests it may be both (and that deeper
analysis is needed to understand the situation better). The larger question
is how to use this kind of information to choose the best locations, over
time, for a diffuse population.

Another natural idea is to co-locate or consolidate conferences.  This is
already standard practice in some communities---e.g., SIGGRAPH, SC, and CVPR
are all ``the'' conference for their areas.  More communities---for example,
the programming languages and security communities, each with four
main conferences every year---could consider following suit.  A different
twist on this idea would be to hold a single conference in two locations
simultaneously (e.g., Boston and Paris), with a video link between the
sites. Switching to virtual program committee meetings for conferences is
another obvious step.\iflater \mwh{data about surprising connection to conference's
  overall footprint?}\bcp{Or leave it for next round of revisions when we
  may have done more/better analysis?}\fi{}  Indeed, this switch is already
happening: we found in a recent poll of ACM SIG chairs that, across ACM, a
majority of the SIGs' main conferences now use online PC meetings rather
than in-person ones.

Another approach is to allow some conference attendees to participate
remotely, saving both emissions and travel time.  Indeed, SIGPLAN already
livestreams all its four main conferences, while SIGACCESS and SIGCHI have
experimented with telepresence robots at their main
conferences~\cite{CHI-remote,Neustaedter}. \iflater\bcp{Next sentence breaks
the flow a bit.  Cut it?}\fi{}Journal-first initiatives, which
make conference attendance optional, can also help.  Looking further
ahead, some communities are already experimenting with fully virtual
% \mwh{reality?}
% BCP: Seems redundant with "fully"
conferences~\cite{OS}.  ACM members can contribute to the
development of technologies for connecting people effectively at a distance.

@@

The problem of climate change is too large and too urgent to leave to world
leaders: organizations at every level must play their part in dealing with
their own contributions, raising awareness, and establishing new ways of
doing business in the lower-carbon future that is coming. For ACM, this
especially means developing strategies for reducing the climate impact of
conference-related travel. You can help in a number of ways: 
\begin{itemize}
\item {\bf Get involved} in raising awareness about this issue in the
organizations you are a part of.  For example, ask the organizers of the ACM
conferences you attend what they are doing about reducing or offsetting
emissions from participants' travel.
\item {\bf Analyze} your organization's carbon-emitting activities to inform a
discussion of how best to reduce its footprint.
\item {\bf Connect} with others that are engaging with these issues---in
your workplace, in your other communities, and within ACM (e.g., on the
acm-climate mailing list~\cite{acm-climate}).
\end{itemize}

@@

About Crista's figure:

@@

Also about Crista's figure:

@@

Also, what about research on telepresence and virtual meetings? There's obviously a lot of work in this area, but we may need more in order to (a) let us effectively trade off options, e.g. virtual PC meeting vs in-person; (b) improve the state of the art to make remote meetings more effective.
Consider also discussing the diversity implications on travel. Reducing travel expectations might make the field more attractive to women, who may face higher social expectations to stay home (and are more likely to be single parents). Surely there is a good article to cite here. I found https://journals.sagepub.com/doi/abs/10.1177/0950017006066999 but maybe there is something better. I'm looking…

@@

Check where we've used ``carbon'' (or CO2) and where ``greenhouse gases''.  Gloss CO2 or CO2e as appropriate.
